---
title: "Relating Lifespan and Immunity"
author: "Mary-Kate"
date: "09/08/2021"
output: 
  workflowr::wflow_html:
    toc: true
    toc_float: true
    code_folding: hide
    fig_caption: true
editor_options:
  chunk_output_type: console
---

# Load Libraries, datastes and then tidy

```{r libraries, message=F, warning=F, results='hide'}
library(survival) # survival functions
library(survminer) # survival functions
library(survivalAnalysis) # survival functions
library(tidyverse) # data tidying
library(tidytidbits) # data tidying
library(DT) # nice data tables
library(plyr) # data import and tidying
library(dplyr) # data tidying
library(coxme) # Cox mixed effect analysis 
library(gtsummary) # Analytic table summaries
library(multcomp) # Linear modeling 
library(scales) # visualization
library(gridExtra) # arrange multiple grid plots
library(grid) # classic graphic functions 
library(ggplot2) # graphic functions
library(kableExtra) # scrolleable tables
library(ggpubr) # level-up plots 
library(purrr) # For mapping through list
library(broom) # Extraction of mapped function calls into dataframes
library(viridis) # colourblind colour palette 
library(cowplot) # Grid plots + labels 
library(rstatix)
library(ggstatsplot) # Adjusted P-values + ggpubr
library(spaMM) # Modelling BLUD
library(data.table)

select = dplyr::select
mutate = dplyr::mutate
Surv = survival::Surv
survfit = survival::survfit
coxme = coxme::coxme

# Font
SuperSmallfont= 6
Smallfont= 10
Mediumfont= 12
Largefont= 14
verySmallfont = 16
pointsize= 0.7
linesize=1.2
meansize = 1.5
Margin=c(0,0,0,0)

fontsizeaxes = 12
fontsizeaxes2 = 10


# DATA import
## 8 representing each of 4 block per time point.
data_path = "data"
files <- list.files(path = data_path, pattern = "^S.*\\.csv$", full.names = T)
files

## Use 'R ldply' to apply read.csv function to all elements and add to dataframe 
data_all <- ldply(files, read_csv)

# DATA tidy
## Select non-empty columns and modify variable classes
data_tidy <-
  data_all %>% 
  select(1:24) %>% 
  mutate(across(where(is.character), as.factor)) %>% 
  mutate(across(Line, as.factor))
         
```

```{r functions, results= FALSE }
# Function to extract CoxME table

 extract_coxme_table <- function (mod){
    beta <- mod[["coefficients"]]
    nvar <- length(beta)
    nfrail <- nrow(mod$var) - nvar
    se <- sqrt(diag(mod$var)[nfrail + 1:nvar])
    z<- round(beta/se, 2)
    p<- format(as.numeric(pchisq((beta/se)^2, 1,lower.tail = F)), 4)
    table=data.frame(cbind(beta,se,z,p))
    return(table)
 }


#Function to extract survival data to plot
ggplotprep2 <- function(x, times){
  d <- data.frame(condition=rep(names(x$strata), x$strata), time=x$time, survival=x$surv, upper=x$upper, lower=x$lower)
  fillup0 <- function(s) rbind(c(condition=s, time=0, survival=1, upper=1, lower=1), d[d$condition==s, ], deparse.level = 0)
  
  indexes <- function(x, time) {
    if(x%in%time) return(x)
    return(floor(time[which.min(abs(time[time<x]-x))]))
  }
  
  fillup <- function(s) {
    d.temp <- d[d$condition==s, ]
    time <- as.numeric(d.temp$time)
    id <- sapply(times, indexes, time=time)
    d.temp <- d.temp[match(id, time), ]
    d.temp$time <- times
    return(d.temp)
  }
  
  if(times[1]==0) d <- do.call("rbind", sapply(names(x$strata), fillup0, simplify=F))
  d <- do.call("rbind", sapply(names(x$strata), fillup, simplify=F))
  clean.name <- function(name) unlist(lapply(strsplit(as.character(name), split="="), function(x) x[2]))
  d <- data.frame(Condition=clean.name(d$condition), Time=as.numeric(d$time), Survival=as.numeric(d$survival), upper=as.numeric(d$upper), lower=as.numeric(d$lower))
  return(d)
}

# Function to generate labels with n numbers
paste_n <- function(y) {
new <-
  n.df %>% 
  filter(Line == y)
new$id <- paste(new$Sex, new$Age, new$n.number, sep = " ")

y_labs <- c(paste(new$id[1]),
      paste(new$id[2]),
      paste(new$id[3]),
      paste(new$id[4]))
y_labs
}

paste_n_block <-function(y,x) {
new <-
  n.df %>% 
  filter(Line == y & Block == x)
new$id <- paste(new$Sex, new$Age, new$n.number, sep = " ")

y_labs <- c(paste(new$id[1]),
      paste(new$id[2]),
      paste(new$id[3]),
      paste(new$id[4]))
y_labs
}
```


# BLUD modelling

BLUD was modelled in each line separately with a) Sex and b) Age as fixed effects, the fit of these models was then used for parametric bootstraping in order to estimate standard errors and coefficients. The first chunk of code below shows the steps for clarity but in interest of time, I will load the ultimate dataframe from these analyses for later use. 

Outliers were removed by subtracting/adding 1.5 times the interquartile range from the 0.25 and 0.75 quantile, respectively.   
-Q2 - 1.5*IQR 
-Q3 + 1.5*IQR

Line 821, the block control, was modeled separately to account for block as a random effect and then combined with the others.

Lines with <5 individuals per condition are annotated with a "*" on the plots.


## Sex Coefficients

Direction and nature of Lifespan was determined through Coxme/coxph on lifespan data I obtained, where food events were censored.  I've included the n number for actual deaths in a table. All coefficients are relative to Females.

My interpretation of these plots is:

* Points in upper left quadrant: Females are more susceptible due to lower tolerance 
* Points in upper right quadrant: Males are more susceptible due to lower resistance
* Points in lower right quadrant: Males are more susceptible due to lower tolerance
* Points in lower left quadrant: Females are more susceptible due to lower resistance

```{r table for numbers of dry deaths}

load("data/lifespanCountdeaths.RData")

lifespan_death_counts <-
   lifespanCountdeaths %>%  
   dplyr::group_by(Line, Sex) %>% 
   dplyr::summarise(mean_deaths = round(mean(Event)))

# Load the beta values for lifespan and merge with death_counts above 
load("data/cox_lifespan_all.RData")

LS_direction <- as.vector(c("Non_dimorphic", "Male_biased", "Non_dimorphic", "Female_biased", "Female_biased", 
                         rep("Male_biased", length.out=7)))
lifespan_beta_counts <- 
   lifespan_death_counts %>% 
   pivot_wider(names_from = Sex, values_from = mean_deaths) %>% 
   merge(cox_lifespan_all, by="Line") %>% 
   dplyr::rename(beta_lifespan = beta,
          se_lifespan = se,
          z_lifespan = z,
          p_lifespan = p,
          Female_deaths = Female,
          Male_deaths = Male) %>% 
   dplyr::mutate(Lifespan_actual = LS_direction)


kable(lifespan_beta_counts[,1:5], caption = "Dry deaths per sex per line and beta estimates") %>% 
  kable_styling(full_width = T)
```

```{r BLUD modelling Sex, warning = FALSE,include=FALSE}

# Try to analyse the raw BLUD data as per David's spaMM method

CONFIG_ <- read_csv("data/COUNTS_config_MK_210628.csv")
CONFIG<-
   CONFIG_ %>%
   select(5,2,3,4,6,7,8,9) %>%
   dplyr::rename(PLATE = Plate)%>% 
   unite(Sample_ID, c(3,5), sep="_", remove = F)
# then assign David's counting for block two, 10 days old
D10B2 <- read.csv("data/BLUD_David_D112.csv")
# finally read in the manual BLUD from day 10, blocks 1 and 3 and pivot_long them
D10B1 <- read.csv("data/BLUD_counts_D111.csv")
D10B3 <- read.csv("data/BLUD_counts_D113.csv")

Block1 <-
   D10B1 %>%
   pivot_longer(cols = contains("0"),
                names_to = "Dilution",
                values_to = "PROVIDENCIA") %>%
   unite(Sample_ID, c(1,3), remove = F) %>%
   separate(Dilution, c("x", "Dilution"), sep = 1) %>%
   mutate(type_of_bacterial_load = paste("BLUD"),
          Volume = 5,
          Antibiotic = paste("Erythromycin")) %>%
select(c(1,3,9,2,11,4,10,7,8))

Block2 <-
   D10B2 %>%
   unite(Sample_ID, c(2,5), remove = F) %>%
   dplyr::rename(type_of_bacterial_load = BLUD_or_SPBL,
                 PLATE = Plate,
                 PROVIDENCIA = BACTERIA) %>%
   mutate(Antibiotic = paste("Erythromycin")) %>% 
   select(c(2,5,1,3,4,6,7:9))

Block3 <-
   D10B3 %>%
   pivot_longer(cols = contains("0"),
                names_to = "Dilution",
                values_to = "PROVIDENCIA") %>%
   select(-1) %>%
   unite(Sample_ID, c(1,3), remove = F) %>%
   separate(Dilution, c("x", "Dilution"), sep = 1) %>%
   mutate(type_of_bacterial_load = paste("BLUD"),
          Volume = 5,
          Antibiotic = paste("Erythromycin")) %>%
   select(c(1,3,9,2,11,4,10,7,8))

# Merge the early counts, do this separately for main counts due to size and related speed
Early_block_counts <- rbind(Block1, Block2, Block3)
All_counts <- rbind(CONFIG,Early_block_counts)

data_tidy_s <-
   data_tidy %>%
   select(1:23) %>%
   unite(Sample_ID, 22:23, sep =  "_" ) %>% 
   dplyr::filter(Sample_ID != "NA_NA")
   


Early_block_counts_dt <- as.data.table(Early_block_counts)
Data_tidy_s_dt <- as.data.table(data_tidy_s)
CONFIG.indiv.ID <- merge(All_counts, data_tidy_s, all.x = T)

Config <-
   CONFIG.indiv.ID %>%
   unite(RepPlateDil,c(4,8), remove = F) %>%
   select(c(1:10,13,14,16,17,18,28)) %>%
   dplyr::group_by(RepPlateDil) %>%
   fill(Lit_lifespan, Line, Sex,Block, Age_at_infection, Time_of_death, .direction = "down")
   

   
#########################################################
# young  
#########################################################

## Find interquatile range so you can remove outliers 1.5*IQR above/below 0.75/0.25 quantile. 
# Range_per_group <- 
#  subset(Config,is.finite(PROVIDENCIA) & PROVIDENCIA !=0 & !is.na(PROVIDENCIA) & !is.na(Line) &
#            Age_at_infection == 10) %>%
#  mutate(Age_at_infection=factor(Age_at_infection),
#         Sex = factor(Sex),
#         Line=factor(Line),
#         across(where(is.character),factor),
#         across(Dilution, as.character),
#         across(Dilution, as.numeric)) %>% 
#   dplyr::group_by(Sample_ID) %>% 
#   mutate(Counts = log(PROVIDENCIA* Dilution)) %>% 
#   dplyr::ungroup() %>% 
#   unite(Group, c(11,13:14), sep = "_", remove = F) %>% 
#   dplyr::group_by(Group) %>% 
#   dplyr::summarise(Q25 = quantile(Counts, probs = 0.25),
  #                  IQR = IQR(Counts),
 #                   Q75 = quantile(Counts, probs = 0.75),
   #                 up = Q75+1.75*IQR,
    #                low = Q25 - 1.5*IQR) 

# dtmp_young = NULL
# dtmp_young =
#  subset(Config,is.finite(PROVIDENCIA) & PROVIDENCIA !=0 & !is.na(PROVIDENCIA) & !is.na(Line) &
#            Age_at_infection == 10) %>%
#  mutate(Age_at_infection=factor(Age_at_infection),
 #        Sex = factor(Sex),
#         Line=factor(Line),
#         across(where(is.character),factor),
#         across(Dilution, as.character),
#         across(Dilution, as.numeric)) %>% 
#   dplyr::group_by(Sample_ID) %>% 
 #  mutate(Counts = log(PROVIDENCIA* Dilution)) %>% 
#   dplyr::ungroup() %>% 
#   unite(Group, c(11,13:14), sep = "_", remove = F) %>% 
#   merge(Range_per_group, by="Group", all = T) %>% 
#   dplyr::group_by(Group) %>% 
#   dplyr::filter(Counts > low & Counts < up)

# Model_lists.1 <-
#   subset(dtmp_young, Line != 821) %>% 
#   dplyr::group_by(Line) %>% 
#   nest() %>% 
#   dplyr::mutate(myoung = map(data, ~fitme(PROVIDENCIA ~ Sex +  (1|Sample_ID) + offset(-log(Dilution)),
#           family=negbin(link="log"),
#           data=.x))) %>% 
#   select(Line, myoung) 

# Block effects in Line 821
# Model_lists <-
#   subset(dtmp_young, Line == 821) %>% 
#   dplyr::group_by(Line) %>% 
#   nest() %>% 
#   dplyr::mutate(myoung = map(data, ~fitme(PROVIDENCIA ~ Sex + (1|Block|Sample_ID) + offset(-log(Dilution)),
#           family=negbin(link="log"),
#           data=.x))) %>% 
#   select(Line, myoung) %>% 
#   rbind(Model_lists.1)

# simPLUD = function(y,what=NULL,m, ...) {
#   data = m$data
#    data$PROVIDENCIA = y ## replaces original response (! more complicated for binomial fits)
#    full_call = getCall(m) ## call for full fit
#    full_call$data = data
#    res = eval(full_call) ## fits the full model on the simulated response
#    if (!is.null(what)) res = eval(what) ## post-process the fit
#    return(res) ## the fit, or anything produced by evaluating 'what'
#  }
  # where the 'what' argument (not required) of myfun() allows one to control
  # what the function returns without redefining the function.

# #################################################################  
# 
##821
# boot.test = NULL
# my821=Model_lists$myoung[[1]]
# boot.test = spaMM_boot(my821, simuland = simPLUD, nsim=50,what=quote(fixef(res)),type='marginal',m=my821)
# save(boot.test,file="output/Bootstrap821_PLUD_Prett_10do_Sex.Rdata")
##304
# boot.young.304 = NULL
# my304=Model_lists$myoung[[5]]
# boot.young.304 = spaMM_boot(my304, simuland = simPLUD, nsim=50,what=quote(fixef(res)),type='marginal', m=my304)
# save(boot.young.304,file="output/Bootstrap304_PLUD_Prett_10do_Sex.Rdata")

# etc..

## Repeated for old

# Ultimately ending up with 
load("data/lPLUD.RData")

####################### my method ###################################

# My own method for calculating BLud

BLUD_mean<- 
   Config %>% 
   filter(!is.na(PROVIDENCIA) & PROVIDENCIA != "Inf") %>% 
   dplyr::mutate(across(Dilution, as.numeric),
                 counts = PROVIDENCIA*Dilution,
          CFU_ml = counts /0.005,
          CFU_FLY = CFU_ml * 0.250) %>% 
   filter(CFU_FLY > 0) %>% 
   dplyr::group_by(Sample_ID) %>% 
   dplyr::summarise(BLUD_mean_reps = mean(CFU_FLY)) 

rest_off_counts <-
   data_tidy %>% 
   filter(BLUD_mean_reps > 0) %>% 
   select(c(1:21,24)) %>% 
   mutate(log_BLUD = log10(BLUD_mean_reps)) %>% 
   unite(Group, c(7,9), sep = "_", remove = F)
# Create sample_ID in data_tidy 
data_tidy_mean <-
   data_tidy %>% 
   select(1:23) %>% 
   unite(Sample_ID, 22:23, sep =  "_" )

surv_blud_merge <- merge(data_tidy_mean,BLUD_mean, by="Sample_ID")


surv_blud <-
   surv_blud_merge %>% 
   select(-1) %>% 
   mutate(log_BLUD = log10(BLUD_mean_reps)) %>% 
   unite(Group, c(7,9), sep = "_", remove = F) %>% 
   rbind(rest_off_counts) %>% 
   mutate(across(Group, as.factor)) %>% 
   unite(Main_Group, c(4,7), remove = F) %>% 
   dplyr::group_by(Main_Group) %>% 
   mutate(Sample_size = length(log_BLUD))
 
   


```

```{r coxph analysis Sex, warning=FALSE, fig.cap="Sex effects on BLUD for female-biased and non-dimorphic lines"}

# For block control, do cox separately and fit block as a random effect:
sex.coxme.age.effs821 <-
   data_tidy %>% 
   dplyr::filter(Line == 821) %>% 
  group_by(Line, Age_at_infection) %>% 
  nest() %>% 
  dplyr::mutate(cox_object = map(data, ~coxme(Surv(time = Time_to_death, event = Censor) ~ Sex + (1|Block),  data=.x))) %>% 
  dplyr::mutate(cox_tidy = map(cox_object, extract_coxme_table)) %>% 
  ungroup() %>% 
  transmute(Line,Age_at_infection, cox_tidy) %>% 
  unnest(cox_tidy) %>% 
   mutate(across(beta, as.numeric))

# Obtain hazard ratios for each line bar 821 at each age
sex.cox.age.effs <-
   data_tidy%>%  
   dplyr::filter(Line != 821) %>% 
  group_by(Line,Age_at_infection) %>% 
  nest() %>% 
  dplyr::mutate(cox_object = map(data, ~coxph(Surv(time = Time_to_death, event = Censor) ~ Sex,  data=.x))) %>% 
  dplyr::mutate(cox_tidy = map(cox_object, extract_coxme_table)) %>% 
  ungroup() %>% 
  transmute(Line,Age_at_infection, cox_tidy) %>% 
  unnest(cox_tidy) %>% 
   mutate(across(beta, as.numeric))

# combine beta coefficients for all lines
total_sex_effs.age <- rbind(sex.coxme.age.effs821, sex.cox.age.effs)

# Create dataframe with line and expected lifespan for faceting
Line_lifespan <- 
   surv_blud %>% 
   select(c(4,6,26)) %>% 
   mutate(Lit_lifespan = recode(Lit_lifespan, FB ="Female_biased", ND ="Non_dimorphic")) %>% 
   separate(Main_Group, c("Line", "Sex", "Age_at_infection")) %>% 
   distinct() %>% 
   pivot_wider(names_from = Sex, values_from = Sample_size) %>% 
   mutate(Female_n = ifelse(Female < 5, paste("*"), paste(" ")),
          Male_n = ifelse(Male < 5, paste("*"), paste(" "))) %>% 
   select(c(1:3, 6:7))

# Now compare lifespan data obtained through censoring of food death

lifespan_death_counts <-
   lifespanCountdeaths %>%  
   dplyr::group_by(Line, Sex) %>% 
   dplyr::summarise(mean_deaths = round(mean(Event)))

# Load the beta values for lifespan and merge with death_counts above 
load("data/cox_lifespan_all.RData")

LS_direction <- as.vector(c("Non_dimorphic", "Male_biased", "Non_dimorphic", "Female_biased", "Female_biased", 
                         rep("Male_biased", length.out=7)))
lifespan_beta_counts <- 
   lifespan_death_counts %>% 
   pivot_wider(names_from = Sex, values_from = mean_deaths) %>% 
   merge(cox_lifespan_all, by="Line") %>% 
   dplyr::rename(beta_lifespan = beta,
          se_lifespan = se,
          z_lifespan = z,
          p_lifespan = p,
          Female_deaths = Female,
          Male_deaths = Male) %>% 
   dplyr::mutate(Lifespan_actual = LS_direction)
                 

# now isolate the mean differences between the log(BLUD) of the sexes for each line

blud_mean_diff <-
   surv_blud %>%   
   select(c(5,9,11,25)) %>% 
   mutate(across(Age_at_infection, factor)) %>% 
   filter(!is.na(log_BLUD)) %>% 
   dplyr::group_by(Line, Sex, Age_at_infection) %>%
   dplyr::summarise(mean_BLUD = mean(log_BLUD)) %>% 
   pivot_wider(names_from = Sex, values_from = mean_BLUD) %>% 
   mutate(SexDiff = Female - Male)

# merge the lplud estimates for confidence intervals
BLUD_error <- merge(blud_mean_diff, lPLUD)

Hazard_Blud <- merge(total_sex_effs.age, merge(BLUD_error, merge(lifespan_beta_counts, Line_lifespan, 
                                               by=c("Line"), all.x = T), by=c("Line", "Age_at_infection")), 
                                               by=c("Line", "Age_at_infection"))
Hazard_Blud$se <- as.numeric(Hazard_Blud$se) 
Hazard_Blud$Age_at_infection <- as.factor(Hazard_Blud$Age_at_infection) 
Hazard_Blud$Lifespan_actual <- as.factor(Hazard_Blud$Lifespan_actual) 

library(ggh4x)
# Female_biased
pFB <- ggplot(subset(Hazard_Blud, Lifespan_actual == "Female_biased"), aes(beta, PLUD), scale="globalminmax") +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = 0, linetype = 2) +
   geom_errorbarh(aes(xmin = beta-se, xmax = beta+se),height =0.2)+
   geom_errorbar(aes(ymax = PLUD_CI_sup, ymin = PLUD_CI_inf),width=0.2)+
   scale_y_continuous("BLUD coeff",
                      limits = c(-1,1.5),
                      breaks=c(seq(-1,1.5,by=1)))+
   scale_x_continuous(limits = c(-1,2),
                      breaks=c(seq(-1,2,by=1)))+
   facet_nested(~Lifespan_actual + Line)+
   geom_point(aes(colour = Age_at_infection), size =2.5) +
   scale_colour_manual("Age",
                       values = c("darkgreen", "darkorchid"))+
   geom_text(aes(label = paste(Female_n)), nudge_y = 0.8,size = 7)+
   geom_text(aes(label = paste(Male_n)), nudge_y = -0.8,size = 7)+
   theme_light()+
   theme(legend.position = "none",
         strip.text.x = element_text(size = 10, face = "bold", colour = "black"),
         aspect.ratio = 1)

# non dimorphic
pND <- ggplot(subset(Hazard_Blud, Lifespan_actual == "Non_dimorphic"), aes(beta, PLUD), scale="globalminmax") +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = 0, linetype = 2) +
   geom_errorbarh(aes(xmin = beta-se, xmax = beta+se),height =0.2)+
   geom_errorbar(aes(ymax = PLUD_CI_sup, ymin = PLUD_CI_inf),width=0.2)+
   scale_y_continuous("BLUD coeff",
                      limits = c(-1,1),
                      breaks=c(seq(-1,1,by=1)))+
   scale_x_continuous(limits = c(-1,2),
                      breaks=c(seq(-1,2,by=1)))+
   facet_nested(~Lifespan_actual+Line)+
   geom_point(aes(colour = Age_at_infection), size =2.5) +
   scale_colour_manual("Age",
                       values = c("darkgreen", "darkorchid"))+
   geom_text(aes(label = paste(Female_n)), nudge_y = 0.8,size = 7)+
   geom_text(aes(label = paste(Male_n)), nudge_y = -0.8,size = 7)+
   theme_light()+
   theme(strip.text.x = element_text(size = 10, face = "bold", colour = "black"),
         aspect.ratio = 1,
         legend.position = "none",
         legend.text = element_text(size = 12, face = "bold"))

# Combine plots

Cox_BLUD_coeff_SEX <- 
  ggpubr::ggarrange(pFB, pND, # list of plots # labels
                  common.legend = T, # COMMON LEGEND, # legend position,
                  legend = "bottom",
                  align = "hv", # Align them both, horizontal and vertical
                  nrow = 1)

Cox_BLUD_coeff_SEX

```

```{r coxph analysis Sex male_biased, warning=FALSE, fig.cap="Sex effects on BLUD for male-biased lines"}
# Male_biased
pMB <-grid.arrange(
   ggplot(subset(Hazard_Blud, Lifespan_actual == "Male_biased" &
                   Line %in% c("730", "821", "386", "907")) , aes(beta, PLUD), scale="globalminmax") +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = 0, linetype = 2) +
   geom_errorbarh(aes(xmin = beta-se, xmax = beta+se),height =0.2)+
   geom_errorbar(aes(ymax = PLUD_CI_sup, ymin = PLUD_CI_inf),width=0.2)+
   scale_y_continuous("BLUD coeff",
                      limits = c(-1,0.5),
                      breaks=c(seq(-0.5,0.5,by=1)))+
   scale_x_continuous(limits = c(-3.5,0.6),
                      breaks=c(seq(-3,1,by=1)))+
   facet_nested(~Lifespan_actual+Line)+
   geom_point(aes(colour = Age_at_infection), size =2.5) +
   scale_colour_manual("Age",
                       values = c("darkgreen", "darkorchid"))+
   geom_text(aes(label = paste(Female_n)), nudge_y = 0.8,size = 7)+
   geom_text(aes(label = paste(Male_n)), nudge_y = -0.8,size = 7)+
   theme_light()+
   theme(strip.text.x = element_text(size = 10, face = "bold", colour = "black"),
         aspect.ratio = 1,
         legend.position = "none",
         legend.text = element_text(size = 12, face = "bold")),
  ggplot(subset(Hazard_Blud, Lifespan_actual == "Male_biased" &
                   Line %in% c("59", "897", "517", "732")) , aes(beta, PLUD), scale="globalminmax") +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = 0, linetype = 2) +
   geom_errorbarh(aes(xmin = beta-se, xmax = beta+se),height =0.2)+
   geom_errorbar(aes(ymax = PLUD_CI_sup, ymin = PLUD_CI_inf),width=0.2)+
   scale_y_continuous("BLUD coeff",
                      limits = c(-1.5,2),
                      breaks=c(seq(-1,2,by=1)))+
   scale_x_continuous(limits = c(-1,2.5),
                      breaks=c(seq(-1,2,by=1)))+
   facet_nested(~Lifespan_actual+Line)+
   geom_point(aes(colour = Age_at_infection), size =2.5) +
   scale_colour_manual("Age",
                       values = c("darkgreen", "darkorchid"))+
   geom_text(aes(label = paste(Female_n)), nudge_y = 0.8,size = 7)+
   geom_text(aes(label = paste(Male_n)), nudge_y = -0.8,size = 7)+
   theme_light()+
   theme(strip.text.x = element_text(size = 10, face = "bold", colour = "black"),
         aspect.ratio = 1,
         legend.position = "none",
         legend.text = element_text(size = 12, face = "bold")),
  nrow=2)



pMB

```

## Age Coeffcients 
All coefficients are relative to young. 

My interpretation of these plots is:

* Points in upper left quadrant: Young individuals are more susceptible due to lower tolerance 
* Points in upper right quadrant: Aged individuals  are more susceptible due to lower resistance
* Points in lower right quadrant: Aged individuals  are more susceptible due to lower tolerance
* Points in lower left quadrant: Young individuals  are more susceptible due to lower resistance

```{r BLUD modelling Age, include=FALSE, warning=FALSE}

load("data/lPLUDage.RData")

# My own method for calculating BLud

BLUD_mean<- 
   Config %>% 
   filter(!is.na(PROVIDENCIA) & PROVIDENCIA != "Inf") %>% 
   dplyr::mutate(across(Dilution, as.numeric),
                 counts = PROVIDENCIA*Dilution,
          CFU_ml = counts /0.005,
          CFU_FLY = CFU_ml * 0.250) %>% 
   filter(CFU_FLY > 0) %>% 
   dplyr::group_by(Sample_ID) %>% 
   dplyr::summarise(BLUD_mean_reps = mean(CFU_FLY)) 

# Create sample_ID in data_tidy 
data_tidy_mean <-
   data_tidy %>% 
   select(1:23) %>% 
   unite(Sample_ID, 22:23, sep =  "_" )

surv_blud_merge <- merge(data_tidy_mean,BLUD_mean, by="Sample_ID")

surv_blud <-
   surv_blud_merge %>% 
   select(-1) %>% 
   mutate(log_BLUD = log10(BLUD_mean_reps)) %>% 
   unite(Group, c(7,9), sep = "_", remove = F) %>% 
   rbind(rest_off_counts) %>% 
   mutate(across(Group, as.factor)) %>% 
   unite(Main_Group, c(4,7), remove = F) %>% 
   dplyr::group_by(Main_Group) %>% 
   mutate(Sample_size = length(log_BLUD))
 


```

```{r Cox analysis Age, warning=FALSE, fig.cap= "Age effects on BLUD for female-biased and non-dimorphic lines"}
# Repeat with model with Age as the term 
# For block control, do cox separately and fit block as a random effect:

data_tidy$Age_at_infection <- as.factor(data_tidy$Age_at_infection)
age.sex.coxme.effs821 <-
   data_tidy %>% 
   dplyr::filter(Line == 821) %>% 
  group_by(Line, Sex) %>% 
  nest() %>% 
  dplyr::mutate(cox_object = map(data, ~coxme(Surv(time = Time_to_death, event = Censor) ~ Age_at_infection + (1|Block),  data=.x))) %>% 
  dplyr::mutate(cox_tidy = map(cox_object, extract_coxme_table)) %>% 
  ungroup() %>% 
  transmute(Line,Sex, cox_tidy) %>% 
  unnest(cox_tidy) %>% 
   mutate(across(beta, as.numeric))

# Obtain hazard ratios for each line bar 821 at each age
age.sex.cox.effs <-
   data_tidy %>%  
   dplyr::filter(Line != 821) %>% 
  group_by(Line,Sex) %>% 
  nest() %>% 
  dplyr::mutate(cox_object = map(data, ~coxph(Surv(time = Time_to_death, event = Censor) ~ Age_at_infection,  data=.x))) %>% 
  dplyr::mutate(cox_tidy = map(cox_object, extract_coxme_table)) %>% 
  ungroup() %>% 
  transmute(Line,Sex, cox_tidy) %>% 
  unnest(cox_tidy) %>% 
   mutate(across(beta, as.numeric))

# combine beta coefficients for all lines
total_agepersex_effs.age <- rbind(age.sex.coxme.effs821, age.sex.cox.effs)

# now isolate the mean differences between the log(BLUD) of young and old for each line

blud_mean_diff <-
   surv_blud %>%   
   select(c(5,9,11,25)) %>% 
   mutate(across(c(Age_at_infection,Sex), factor)) %>% 
   filter(!is.na(log_BLUD)) %>% 
   dplyr::group_by(Line, Sex, Age_at_infection) %>%
   dplyr::summarise(mean_BLUD = mean(log_BLUD)) %>% 
   mutate(Age_at_infection = recode(Age_at_infection, '10' = "Young", '35' = "Old")) %>% 
   pivot_wider(names_from = Age_at_infection, values_from = mean_BLUD) %>% 
   mutate(SexDiff = Young - Old)


Line_lifespan <- 
   surv_blud %>% 
   select(c(4,6,26)) %>% 
   mutate(Lit_lifespan = recode(Lit_lifespan, FB ="Female_biased", ND ="Non_dimorphic")) %>% 
   separate(Main_Group, c("Line", "Sex", "Age_at_infection")) %>% 
   distinct() %>% 
   mutate(Age_at_infection = recode(Age_at_infection, '10' = "Young", '35' = "Old")) %>% 
   pivot_wider(names_from = Age_at_infection, values_from = Sample_size) %>% 
   mutate(Young_n = ifelse(Young <5, paste("*"), paste(" ")),
          Old_n = ifelse(Old < 5, paste("*"), paste(" "))) %>% 
   select(c(1:3, 6:7))


BLUD_error <- merge(blud_mean_diff, lPLUDage)

Hazard_Blud <- merge(total_agepersex_effs.age, merge(BLUD_error, merge(lifespan_beta_counts, Line_lifespan, 
                                               by=c("Line"), all.x = T), by=c("Line", "Sex")), 
                                               by=c("Line", "Sex"))
Hazard_Blud$se <- as.numeric(Hazard_Blud$se) 
Hazard_Blud$Sex<- as.factor(Hazard_Blud$Sex) 
Hazard_Blud$Lifespan_actual <- as.factor(Hazard_Blud$Lifespan_actual) 


library(ggh4x)

# Female_biased
pFB <- ggplot(subset(Hazard_Blud, Lifespan_actual == "Female_biased"), aes(beta, PLUD), scale="globalminmax") +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = 0, linetype = 2) +
   geom_errorbarh(aes(xmin = beta-se, xmax = beta+se),height =0.2)+
   geom_errorbar(aes(ymax = PLUD_CI_sup, ymin = PLUD_CI_inf),width=0.2)+
   scale_y_continuous("BLUD coeff",
                      limits = c(-1.5,0.5),
                      breaks=c(seq(-1.5,0.5,by=1)))+
   scale_x_continuous(limits = c(-1,2),
                      breaks=c(seq(-1,2,by=1)))+
   facet_nested(~Lifespan_actual+Line)+
   geom_point(aes(colour = Sex), size =2.5) +
   scale_colour_manual(values = c("firebrick", "deepskyblue1"))+
   geom_text(aes(label = paste(Young_n)), nudge_y = 1.5,size = 7)+
   geom_text(aes(label = paste(Old_n)), nudge_y = -1,size = 7)+
   theme_light()+
   theme(legend.position = "none",
         strip.text.x = element_text(size = 10, face = "bold", colour = "black"),
         aspect.ratio = 1)

# non dimorphic
pND <- 
  ggplot(subset(Hazard_Blud, Lifespan_actual == "Non_dimorphic"), aes(beta, PLUD), scale="globalminmax") +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = 0, linetype = 2) +
   geom_errorbarh(aes(xmin = beta-se, xmax = beta+se),height =0.2)+
   geom_errorbar(aes(ymax = PLUD_CI_sup, ymin = PLUD_CI_inf),width=0.2)+
   scale_y_continuous("BLUD coeff",
                      limits = c(-1.5,2.5),
                      breaks=c(seq(-1.5,1.5,by=1)))+
   scale_x_continuous(limits = c(-1,1),
                      breaks=c(seq(-1,1,by=1)))+
   facet_nested(~Lifespan_actual+Line)+
   geom_point(aes(colour = Sex), size =2.5) +
   scale_colour_manual(values = c("firebrick", "deepskyblue1"))+
   geom_text(aes(label = paste(Young_n)), nudge_y = 1.5,size = 7)+
   geom_text(aes(label = paste(Old_n)), nudge_y = -1,size = 7)+
   theme_light()+
   theme(strip.text.x = element_text(size = 10, face = "bold", colour = "black"),
         aspect.ratio = 1,
         legend.position = "none",
         legend.text = element_text(size = 12, face = "bold"))

Cox_BLUD_coeff_AGE <- 
  ggpubr::ggarrange(pFB, pND, 
                    common.legend = T,
                    legend = "bottom",
                  align = "hv", # Align them both, horizontal and vertical
                  nrow = 1)
Cox_BLUD_coeff_AGE
   


```

```{r coxph analysis Age male_biased, warning=FALSE, fig.cap="Age effects on BLUD for male-biased lines"}

pMB <- grid.arrange(
  ggplot(subset(Hazard_Blud, Lifespan_actual == "Male_biased" &
                   Line %in% c("732", "821", "59", "386")), aes(beta, PLUD), scale="globalminmax") +
  geom_hline(yintercept = 0, linetype = 2) +
   geom_errorbarh(aes(xmin = beta-se, xmax = beta+se),height =0.2)+
   geom_errorbar(aes(ymax = PLUD_CI_sup, ymin = PLUD_CI_inf),width=0.2)+
   scale_y_continuous("BLUD coeff",
                      limits = c(-1,1.5),
                      breaks=c(seq(-1,1,by=1)))+
   scale_x_continuous(limits = c(0,2),
                      breaks=c(seq(0,2,by=1)))+
   facet_nested(~Lifespan_actual+Line)+
   geom_point(aes(colour = Sex), size =2.5) +
   scale_colour_manual(values = c("firebrick", "deepskyblue1"))+
   geom_text(aes(label = paste(Young_n)), nudge_y = 1.5,size = 7)+
   geom_text(aes(label = paste(Old_n)), nudge_y = -1,size = 7)+
   theme_light()+
   theme(strip.text.x = element_text(size = 10, face = "bold", colour = "black"),
         aspect.ratio = 1,
         legend.position = "none",
         legend.text = element_text(size = 12, face = "bold")),
  ggplot(subset(Hazard_Blud, Lifespan_actual == "Male_biased" &
                   Line %in% c("897", "907", "517", "730")), aes(beta, PLUD), scale="globalminmax") +
  geom_hline(yintercept = 0, linetype = 2) +
   geom_errorbarh(aes(xmin = beta-se, xmax = beta+se),height =0.2)+
   geom_errorbar(aes(ymax = PLUD_CI_sup, ymin = PLUD_CI_inf),width=0.2)+
   scale_y_continuous("BLUD coeff",
                      limits = c(-2,1.5),
                      breaks=c(seq(-2,1,by=1)))+
   scale_x_continuous(limits = c(0,4.2),
                      breaks=c(seq(0,4,by=1)))+
   facet_nested(~Lifespan_actual+Line)+
   geom_point(aes(colour = Sex), size =2.5) +
   scale_colour_manual(values = c("firebrick", "deepskyblue1"))+
   geom_text(aes(label = paste(Young_n)), nudge_y = 1.5,size = 7)+
   geom_text(aes(label = paste(Old_n)), nudge_y = -1,size = 7)+
   theme_light()+
   theme(strip.text.x = element_text(size = 10, face = "bold", colour = "black"),
         aspect.ratio = 1,
         legend.position = "none",
         legend.text = element_text(size = 12, face = "bold")),
  nrow=2)
pMB


```


# Age-effects with all lines plotted as replicates 

* An unsurprising effect of age in all cases.
* Older males of male-biased lines seem to be more neatly divided into lower tolerance or lower resistance whereas females are a mixture of both.


```{r Age beta and coeffiecients for all lines as replicates, warning=FALSE, echo=FALSE, fig.cap="Age effects on BLUD for all lines"}

All_lines_age <- ggplot(Hazard_Blud, aes(beta, PLUD), scale="globalminmax") +
   facet_nested(Sex ~ Lifespan_actual)+
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = 0, linetype = 2) +
   geom_errorbarh(aes(xmin = beta-se, xmax = beta+se),height =0.2)+
   geom_errorbar(aes(ymax = PLUD_CI_sup, ymin = PLUD_CI_inf),width=0.2)+
   scale_y_continuous("BLUD coeff",
                      limits = c(-2,2.5),
                      breaks=c(seq(-2,2.5,by=1)))+
   scale_x_continuous(limits = c(-1,4),
                      breaks=c(seq(-1,3,by=1)))+
   geom_point(aes(colour = Sex), size =2.5) +
   scale_colour_manual("Sex",
                       values = c("firebrick", "deepskyblue1"))+
   geom_text(aes(label = paste(Young_n)), nudge_y = 1.5,size = 7)+
   geom_text(aes(label = paste(Old_n)), nudge_y = -1,size = 7)+
   theme_light()+
   theme(legend.position = "bottom",
         strip.text.x = element_text(size = 10, face = "bold", colour = "black"),
         aspect.ratio = 1)
All_lines_age


```

```{r Age beta and coeffiecients for all lines ungrouped as replicates, warning=FALSE, echo=FALSE, fig.cap="Age effects on BLUD for all lines, ungrouped"}

ungrouped_All_lines_age <- 
  ggplot(Hazard_Blud, aes(beta, PLUD), scale="globalminmax") +
   facet_nested(~Sex)+
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = 0, linetype = 2) +
   geom_errorbarh(aes(xmin = beta-se, xmax = beta+se),height =0.2)+
   geom_errorbar(aes(ymax = PLUD_CI_sup, ymin = PLUD_CI_inf),width=0.2)+
   scale_y_continuous("BLUD coeff",
                      limits = c(-2,2.5),
                      breaks=c(seq(-2,2.5,by=1)))+
   scale_x_continuous(limits = c(-1,4),
                      breaks=c(seq(-1,3,by=1)))+
   geom_point(aes(colour = Sex), size =2.5) +
   scale_colour_manual("Sex",
                       values = c("firebrick", "deepskyblue1"))+
   geom_text(aes(label = paste(Young_n)), nudge_y = 1.5,size = 7)+
   geom_text(aes(label = paste(Old_n)), nudge_y = -1,size = 7)+
   theme_light()+
   theme(legend.position = "bottom",
         strip.text.x = element_text(size = 10, face = "bold", colour = "black"),
         aspect.ratio = 1)
ungrouped_All_lines_age


```

# Sex-effects with all lines plotted as replicates 

* Males are more susceptible at 10days old in female-biased lines, but this is seen also in non-dimorphic lines.
* Trends are mixed in male-biased lines.
* Effects tend to be greater at 10days old, but still see a spread in survival among male-biased lines. 

```{r Sex beta and coeffiecients for all lines as replicates, echo=FALSE, warning=FALSE, fig.cap="Sex effects on BLUD for all lines"}

total_sex_effs.age <- rbind(sex.coxme.age.effs821, sex.cox.age.effs)

# Create dataframe with line and expected lifespan for faceting
Line_lifespan <- 
   surv_blud %>% 
   select(c(4,6,26)) %>% 
   mutate(Lit_lifespan = recode(Lit_lifespan, FB ="Female_biased", ND ="Non_dimorphic")) %>% 
   separate(Main_Group, c("Line", "Sex", "Age_at_infection")) %>% 
   distinct() %>% 
   pivot_wider(names_from = Sex, values_from = Sample_size) %>% 
   mutate(Female_n = ifelse(Female < 5, paste("*"), paste(" ")),
          Male_n = ifelse(Male < 5, paste("*"), paste(" "))) %>% 
   select(c(1:3, 6:7))

blud_mean_diff <-
   surv_blud %>%   
   select(c(5,9,11,25)) %>% 
   mutate(across(Age_at_infection, factor)) %>% 
   filter(!is.na(log_BLUD)) %>% 
   dplyr::group_by(Line, Sex, Age_at_infection) %>%
   dplyr::summarise(mean_BLUD = mean(log_BLUD)) %>% 
   pivot_wider(names_from = Sex, values_from = mean_BLUD) %>% 
   mutate(SexDiff = Female - Male)

# merge the lplud estimates for confidence intervals
BLUD_error <- merge(blud_mean_diff, lPLUD)

Hazard_Blud <- merge(total_sex_effs.age, merge(BLUD_error, merge(lifespan_beta_counts, Line_lifespan, 
                                               by=c("Line"), all.x = T), by=c("Line", "Age_at_infection")), 
                                               by=c("Line", "Age_at_infection"))
Hazard_Blud$se <- as.numeric(Hazard_Blud$se) 
Hazard_Blud$Age_at_infection <- as.factor(Hazard_Blud$Age_at_infection) 
Hazard_Blud$Lifespan_actual <- as.factor(Hazard_Blud$Lifespan_actual)

All_lines_sex <- ggplot(Hazard_Blud, aes(beta, PLUD), scale="globalminmax") +
   facet_nested(Age_at_infection ~ Lifespan_actual)+
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = 0, linetype = 2) +
   geom_errorbarh(aes(xmin = beta-se, xmax = beta+se),height =0.2)+
   geom_errorbar(aes(ymax = PLUD_CI_sup, ymin = PLUD_CI_inf),width=0.2)+
   scale_y_continuous("BLUD coeff",
                      limits = c(-2,2),
                      breaks=c(seq(-2,2,by=1)))+
   scale_x_continuous(limits = c(-2,2),
                      breaks=c(seq(-2,2,by=1)))+
   geom_point(aes(colour = Age_at_infection), size =2.5) +
   scale_colour_manual("Age",
                       values = c("darkgreen", "darkorchid"))+
   geom_text(aes(label = paste(Female_n)), nudge_y = 0.8,size = 7)+
   geom_text(aes(label = paste(Male_n)), nudge_y = -0.8,size = 7)+
   theme_light()+
   theme(legend.position = "bottom",
         strip.text.x = element_text(size = 10, face = "bold", colour = "black"),
         aspect.ratio = 1)
All_lines_sex

```

```{r Sex beta and coeffiecients for all lines ungrouped as replicates, echo=FALSE, warning=FALSE, fig.cap="Sex effects on BLUD for all lines, ungrouped"}
ungrouped_All_lines_sex <- 
  ggplot(Hazard_Blud, aes(beta, PLUD), scale="globalminmax") +
   facet_wrap(Age_at_infection ~ .)+
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = 0, linetype = 2) +
   geom_errorbarh(aes(xmin = beta-se, xmax = beta+se),height =0.2)+
   geom_errorbar(aes(ymax = PLUD_CI_sup, ymin = PLUD_CI_inf),width=0.2)+
   scale_y_continuous("BLUD coeff",
                      limits = c(-2,2),
                      breaks=c(seq(-2,2,by=1)))+
   scale_x_continuous(limits = c(-2,2),
                      breaks=c(seq(-2,2,by=1)))+
   geom_point(aes(colour = Age_at_infection), size =2.5) +
   scale_colour_manual("Age",
                       values = c("darkgreen", "darkorchid"))+
   geom_text(aes(label = paste(Female_n)), nudge_y = 1.2,size = 7)+
   geom_text(aes(label = paste(Male_n)), nudge_y = -0.8,size = 7)+
   theme_light()+
   theme(legend.position = "bottom",
         strip.text.x = element_text(size = 10, face = "bold", colour = "black"),
         aspect.ratio = 1)
ungrouped_All_lines_sex

```
